Automated discovery \\ of interpretable gravitational-wave population models

### 2022-07-11 src/tex/ms.tex
Additional text cleanup
Inline score definition to get more space

Internal to \pysr, a multi-objective optimization problem is solved: the evolution seeks to find expressions which are as accurate as possible, while being as simple as possible.
Formally, for our setting, \pysr solves the following optimization problem:
\begin{align}
    s_c &= 
        \mathop{\arg\min}_{
            \tiny
            \begin{array}{c}s\in\mathcal{S}\\ \abs{s} = c \end{array}
        }
        \mathcal{L}(s),
    \quad\text{for each } c\in \mathcal{C},\nonumber
\end{align}
where the loss function $\mathcal{L}$ is defined as:
\begin{equation}
\mathcal{L}(s) = \sum_i \ell(s(\mathbf{x}_i), y_i),
\end{equation}
where $s$ is an expression, 
$\mathcal{S}$ is a space of possible expressions considered,
$\abs{s}$ is the complexity of an expression $s$,
$\ell$ is the loss function, and $\{(x_i, y_i)\}$ is the dataset
for $x_i$ each input and $y_i$ the label.
The most accurate expression at each $c$, denoted as $s_c$, is computed for a given set range of $c$, which we choose as $c\in\mathcal{C}\equiv \{1, \ldots, 50\}$.
Complexity is defined here
as the number of symbols in the expression (including repeats),
and $\ell$ is defined as $(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.

Next, given the set of the best expressions at each complexity $\mathcal{S}^\ast = \{s_c\}_{c\in\{1,\ldots,80\}}$,
Next, we can construct the \textit{Pareto front}, defined as:
\begin{equation}
    \hat{\mathcal{S}}^\ast = \{s_c \ |\ \mathcal{L}(s_c) < \mathcal{L}(s_q) \forall q < c \},
\end{equation}
such that each remaining expression is more accurate than all simpler expressions.
Using this, we can define the \textit{score} of an expression as the slope of the Pareto front, equal to:
\begin{equation}
    \text{score} = \frac{\Delta (-\log \mathcal{L}(s))}{\Delta(\abs{s})},
\end{equation}
which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \citep[see, e.g.,][]{schmidtDistillingFreeFormNatural2009,cranmerDiscoveringSymbolicModels2020}.

@@

In order to maximally extract information from the GW catalog, the community has employed Hierarchical Bayesian Analysis (HBA) to analyze the distribution of GW events.
HBA takes advantage of the fact that measurement uncertainties for each event are conditionally independent from the population model, so we can deconvolve the measurement uncertainty for each event (\cite{10.1214/10-AOAS439} describes this as extreme deconvolution) while fitting for the underlying distribution of GW events\footnote{For interested readers, \cite{Mandel:2018mve,Vitale:2020aaz,2019MNRAS.484.4008G} are excellent references with practical examples on how to apply the HBA framework to GW data.
}.

>>>>

Internal to \pysr, a multi-objective optimization problem is solved: the evolution seeks to find expressions which are as accurate as possible, while being as simple as possible.
Formally, for our setting, \pysr solves the following optimization problem:
\begin{align}
    s_c &= 
        \mathop{\arg\min}_{
            \tiny
            \begin{array}{c}s\in\mathcal{S}\\ \abs{s} = c \end{array}
        }
        \mathcal{L}(s),
    \quad\text{for each } c\in \mathcal{C},
\end{align}
for the loss function $\mathcal{L}(s) = \sum_i \ell(s(\mathbf{x}_i), y_i),$
where $s$ is an expression, 
$\mathcal{S}$ is a space of possible expressions considered,
$\abs{s}$ is the complexity of an expression $s$,
$\ell$ is the loss function, and $\{(x_i, y_i)\}$ is the dataset
for $x_i$ each input and $y_i$ the label.
The most accurate expression at each $c$, denoted as $s_c$, is computed for a given set range of $c$, which we choose as $c\in\mathcal{C}\equiv \{1, \ldots, 50\}$.
Complexity is defined here
as the number of symbols in the expression (including repeats),
and $\ell$ is defined as $(\hat{y}_i, y_i) = \qty(\hat{y}_i - y_i)^2$.

Next, given the set of the best expressions at each complexity $\mathcal{S}^\ast = \{s_c\}_{c\in\{1,\ldots,80\}}$,
Next, we can construct the \textit{Pareto front}, defined as:
\begin{equation}
    \hat{\mathcal{S}}^\ast = \{s_c \ |\ \mathcal{L}(s_c) < \mathcal{L}(s_q) \forall q < c \},
\end{equation}
such that each remaining expression is more accurate than all simpler expressions.
Using this, we can define the \textit{score} of an expression as the slope of the Pareto front, equal to: $\text{score} = -\Delta(\log \mathcal{L}(s_c))/\Delta c,$
which is a heuristic for the quality of a particular equation, used in the symbolic regression literature \citep[see, e.g.,][]{schmidtDistillingFreeFormNatural2009,cranmerDiscoveringSymbolicModels2020}.

@@

In order to maximally extract information from the GW catalog, the community has employed Hierarchical Bayesian Analysis (HBA) to analyze the distribution of GW events.
HBA takes advantage of the fact that measurement uncertainties for each event are conditionally independent from the population model, so we can deconvolve the measurement uncertainty for each event (\citealt{10.1214/10-AOAS439} describes this as extreme deconvolution) while fitting for the underlying distribution of GW events\footnote{For interested readers, \cite{Mandel:2018mve,Vitale:2020aaz,2019MNRAS.484.4008G} are excellent references with practical examples on how to apply the HBA framework to GW data.
}.
